% vim: spell spelllang=en_gb
\chapter{Introduction}\label{sec:introduction}

Earlier in this century, floods around Lake Vänern and Arvika have costed Sweden an estimate of 11.1
billion Swedish Krona for damages and repairs \cite{RiverFloodsSweden2022}. Counties of Dalarna and
Gävleborg suffered from flash floods in 2021, disturbing the daily life of their citizens and
damaging public and private properties \cite{daviesSwedenFlashFloods2021}. Flooding is a devastating
natural disaster that threatens the livelihood of people and the infrastructure of communities
around the world \cite{Floodlist2021}.

To facilitate the process of emergency management during these hazardous events, early warning
systems analyse their risk, monitor and warn the public while ensuring their readiness
\cite{contributorsEarlyWarningSystem2022}. Traditionally, meteorologists forecast the weather by
relying on tools such as gauges, satellites, and radars for data extraction. The emergence of social
media platforms such as Twitter provide individuals with a public space to share their experience,
effectively creating another potential data.

Researchers started harnessing this new wealth of information to aid the disaster management
procedure. Twitter streaming \ac{API} makes it possible to create a monitoring system for early
event detection on a global \cite{debruijnGlobalDatabaseHistoric2019b} and local
\cite{barkerDevelopmentNationalscaleRealtime2019} scales. Another use for it would be identifying
victims in real-time, locating their physical location, and communicating the information to rescue
teams \cite{singhEventClassificationLocation2019}. After the threat subsides, emergency managers can
use relevant tweets to assess the impact and plan the recovery phase
\cite{barkerDevelopmentNationalscaleRealtime2019}. To prepare for future floods, authoritative
entities can make informed actions by analysing historical data and determining the locations
suffering from recurrent calamities. This newly acquired knowledge can augment weather warning
systems' pipelines improving their accuracies such as \ac{SMHI}, a Swedish expert authority with a
global perspective and a vital task in predicting changes in weather, water, and climate
\cite{SMHI2021}.

Most research addresses the problem on a global or national scale, and none addresses Sweden
specifically. This thesis project covers this gap by using Swedish tweets to extract relevant
knowledge about flood events in Sweden with the focus on answering the following questions:
\begin{enumerate} 
  \item What methods can be used to classify Swedish tweets related to flood events?
  \item How to extract the locations mentioned in the tweets? 
  \item What insights can be extracted by using text analysis on the tweets?
  \item What visualizations can be used to represent the results? 
\end{enumerate}

The project focuses on providing a proof of concept for addressing the questions above, and it will
not design a solution to make a production-ready product; i.e., the engineering challenges, such as
automation, scalability, and ease of deployment, are a delimitation. 

This thesis project is a part of a research project, AI for Climate Adaptation
\cite{nesetAI4ClimateAdaptation} at \ac{LiU} in collaboration with the \ac{SMHI}. It implements a
pipeline that provides a visual representation of tweets related to flood events in Sweden. First,
relevant tweets are pulled, processed, and classified from the Twitter API using data mining
techniques. Second, physical locations are extracted from tweets mentioning flood events employing
\ac{NER} and gazetteers. Finally, the results are presented on a spatiotemporal visualization, and
text analysis techniques are done on the tweets. For verification purposes, the pipeline is applied
on a week's worth of tweets after past flood events.
